Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-03-10 supplement.tex
Update supplement.tex

\begin{equation}
	\Delta^2=\mathrm{E}\left\{\left(\bar{X}_1-\bar{X}_2\right)^2\right\}+\mathrm{E}\left\{V_1\right\}+\mathrm{E}\left\{V_2\right\}
\end{equation}

where $\bar{X}_1$ and $V_1$ and $\bar{X}_2$ and $V_2$ are measurement means and variances for the populations from which fossil 1 and fossil 2 are sampled. If we assume that the trait being measured is evolving neutrally, which seems reasonable as an approximation based on studies of morphological (cranial) variation in humans and Neanderthals \tdwcomment{cite here: main text references 36,37, and 38}, then

@@

\begin{equation}
\mathrm{E}\left\{V_B\right\}=\frac{1}{h^2}\sigma_m^2\tau_{W_2}
\end{equation}

where $\sigma^2_m$ is the additive genetic variance introduced by mutation (per zygote per generation), $h^2$ is the (narrow sense) heritability, $\tau_{W_1}$ and $\tau_{W_2}$ are the average coalescence time of pairs of alleles from the same population and $\tau_B$ is the average coalescence time of pairs of alleles from different populations \tdwcomment{cite here: Whitlock 1999 doi.org/10.1017/S0016672399004127 and Weaver 2016 doi.org/10.1002/ece3.2522}. Substituting and simplifying gives

\begin{equation}
	\Delta^2=\sigma_m^2 \left( 4\tau_B+\frac{1-2h^2}{h^2} \left( \tau_{W_1}+\tau_{W_2} \right) \right)
\end{equation}

which relates $\Delta^2$ to between and within population coalescence times. As a further approximation, if we assume that the heritability of the measurement is close to one half, then

\begin{equation}
	\Delta^2\approx4\sigma_m^2\tau_B
\end{equation}

which shows that under these assumptions $\Delta^2$ is proportional to $\tau_B$.

>>>>

\begin{equation}
	\Delta^2=\mathrm{E}\left\{\left(\bar{X}_1-\bar{X}_2\right)^2\right\}+\mathrm{E}\left\{V_1\right\}+\mathrm{E}\left\{V_2\right\}
\end{equation}
where $\bar{X}_1$ and $V_1$ and $\bar{X}_2$ and $V_2$ are measurement means and variances for the populations from which fossil 1 and fossil 2 are sampled. If we assume that the trait being measured is evolving neutrally, which seems reasonable as an approximation based on studies of morphological (cranial) variation in humans and Neanderthals \tdwcomment{cite here: main text references 36,37, and 38}, then

@@

\begin{equation}
\mathrm{E}\left\{V_B\right\}=\frac{1}{h^2}\sigma_m^2\tau_{W_2}
\end{equation}
where $\sigma^2_m$ is the additive genetic variance introduced by mutation (per zygote per generation), $h^2$ is the (narrow sense) heritability, $\tau_{W_1}$ and $\tau_{W_2}$ are the average coalescence time of pairs of alleles from the same population and $\tau_B$ is the average coalescence time of pairs of alleles from different populations \tdwcomment{cite here: Whitlock 1999 doi.org/10.1017/S0016672399004127 and Weaver 2016 doi.org/10.1002/ece3.2522}. Substituting and simplifying gives

\begin{equation}
	\Delta^2=\sigma_m^2 \left( 4\tau_B+\frac{1-2h^2}{h^2} \left( \tau_{W_1}+\tau_{W_2} \right) \right)
\end{equation}
which relates $\Delta^2$ to between and within population coalescence times. As a further approximation, if we assume that the heritability of the measurement is close to one half, then

\begin{equation}
	\Delta^2\approx4\sigma_m^2\tau_B
\end{equation}
which shows that under these assumptions $\Delta^2$ is proportional to $\tau_B$.
